---
title: "Flurex Microbiome Beta Diversity"
author: "Claus Lykkebo, Martin Steen Mortensen"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    toc_depth: 4
    collapsed: false
    code_folding: hide
    number_sections: true
knit: (function(inputFile, encoding) { 
    rmarkdown::render(
        inputFile, encoding = encoding,
        output_dir = file.path(dirname(inputFile),"output"),
        output_file = paste0("Flurex_", Sys.Date(), "_BetaDiversity.html")) 
    })
params:
    input: "R_objects/Phyloseq_15000.Rdata"
    batch: "run"
    metrices: ["UniF","Bray"] # To run weighted unifrac and jaccard add the following to brackets: "WUF","Jaccard"
---

# BETA DIVERSITY

This Rmarkdown contains the commands necessary to perform beta diversity analysis of the output from the DF_GMH_PIPELINE. It is expected that the data has been imported, cleaned, and saved following the script 1_Import_QC.Rmd prior to using this script. I recommend visiting the ["Analysis of community ecology data in R"](https://www.davidzeleny.net/anadat-r/doku.php/en:start) to read about the theory behind the alpha and beta diversity and examples of the necessary R-code to execute them. Other excellent source of help is the R [cheat-sheets](https://www.rstudio.com/resources/cheatsheets/) and for problems related to Rmarkdown I suggest this [online Book](https://bookdown.org/yihui/rmarkdown/).

**Beta diversity**, also called "between sample diversity" is a measurement of the distance, or difference, between samples. First step will be to calculate the beta diversity, second to identify batch effects, and lastly to determine project effects

# INFO

This document has been fitted to perform beta diversity calculations of phyloseq element from the Import & QC script. It has been fitted for data from Flurex (internal project name: R20-22). The project data contains 16S V3 amplicon sequencing from feces samples from 4 different days and cecum and ileum samples from one day from 48 rats. There are feces samples from all rats from day 0, 2, 4 and 8. Cecum and ileum samples were all collected at dissection on day 8. Please read the method section for this paper for details on experimental setup.

Metadata contains sample information regarding: 
-   rat number/name, cage number (rats were co-caged 2-2), day of sampling, material (feces, cecum, ileum), sequencing information (gram of material used for extracted DNA, DNA concentrations, primers, barcodes for demultiplexing, batch number ("run" = c2, c3, c4)), treatment groups. Several columns have been created for easy sample grouping during analysis, including for distinguishing between vancomycin and PFOS treatment alone.

Metadata contains results from:
-   Bodyweight per sampling day, while cecum and liver weight for day 8 (addition column with organ weights normalised to bodyweight).

-   PFOS quantification of:
    +   Blood serum from day 4 and 8 measured in µg/mL
    +   Liver tissue from day 8 measured in mg PFOS in full liver weight
-   Short-chain fatty acids quantification of 10 compounds in colonic water given in mM from day 8: acetic acid, formic acid, propanoic acid, 2methyl-propanoic acid, butanoic acid, 3methyl-butanoic acid, pentanoic acid, 4methyl-pentanoic acid, hexanoic acid and heptanoic acid


```{r setup, eval=TRUE, echo=TRUE, message=FALSE,warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(tidyverse)
library(phyloseq)
library(ggpubr)
library(rstatix)
library(vegan)
library(ape)
library(kableExtra)
library(reshape2)
library(dplyr)
library(ggplot2)
library(cowplot)
library(ggExtra)

# Create used folders if missing
if (!file.exists("R_objects")) dir.create(file.path(getwd(), "R_objects"))
if (!file.exists("plots")) dir.create(file.path(getwd(), "plots"))
if (!file.exists("plots/bdiv")) dir.create(file.path(getwd(), "plots/bdiv"))
if (!file.exists("tables")) dir.create(file.path(getwd(), "tables"))
if (!file.exists("scripts")) dir.create(file.path(getwd(), "scripts"))

# Save params
saveRDS(params, file = "R_Objects/bdiv_params.RDS")
```

## SCRIPTS {.tabset .tabset-fade .tabset-pills}


This section contains the scripts for the custom functions used in this pipeline

### MULTIPLE RAREFY

Here is a function that rarefies each sample of a phyloseq object, calculates the distance between them, and chooses the most central as a representative sample for any following analyses.

```{r multiple_rarefy_function, eval=TRUE}
multiple_rarefy <- function(physeq, ntables=100, depth = min(rowSums(rawtab))*0.9, distmethod="bray", summarymeasure=mean, seedstart=500, verbose=TRUE) {
  require("vegan")
  # Orientate the OTU correctly
  if (taxa_are_rows(physeq)){rawtab<-unclass(t(otu_table(physeq)))} else rawtab <- unclass(otu_table(physeq))
  
  # Ignore samples below rarefaction depth
  ind <- (rowSums(rawtab) < depth)
  sam.discard <- rownames(rawtab)[ind]
  otu.tab <- rawtab[!ind, ]
  
  # Rarefaction function
  rarefy <- function(x, depth) {
    y <- sample(rep(1:length(x), x), depth)
    y.tab <- table(y)
    j <- numeric(length(x))
    j[as.numeric(names(y.tab))] <- y.tab
    j
  }
  
  # Table to output rarefied data
  final_tab = c()
  
  # Run each sample separately
  for (z in 1:nrow(otu.tab)) {
    if (verbose==TRUE) {
      print(paste("Rarefaction sample number", z, sep=" "))
    }
    numbers <- otu.tab[z,]
    
    # Rarefy the sample ntables times
    set.seed(seedstart + z)
    rare_tab <- lapply(1:ntables,function(k) rarefy(numbers,depth))
    
    rare_tab <- do.call(rbind, rare_tab)
    # # Remove columns with no reads
    # rare_tab_no_zero <- rare_tab[,colSums(rare_tab) != 0]
    # # distance across reps for subject z
    distmat = as.matrix(vegdist(rare_tab, method=distmethod)) 
    # calculate mean distance for each rep 
    distsummary = apply(distmat, 2, summarymeasure)
    # the best rep is the one with the mean distance to all other reps. (in case of ties, just select the first)
    whichbestrep = which(distsummary == min(distsummary))[1]  
    # select that rep only for subject z
    bestrep = rare_tab[whichbestrep,]
    # build that rep for subject y into final table
    final_tab = rbind(final_tab, bestrep) 
  }
  
  # Remove samples with too few reads
  physeq <- prune_samples(!sample_names(physeq) %in% sam.discard, physeq) 
  # Reformat final tab and return to the physeq object
  rownames(final_tab) = rownames(otu.tab)
  colnames(final_tab) = colnames(otu.tab)
  otu_table(physeq) <- otu_table(t(final_tab), taxa_are_rows = T)
  
  # Return physeq to the environment
  return(physeq)
}

# save functions
save(multiple_rarefy, file = "scripts/mrarefy.Rdata")

# clear the environment and release memory
rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
invisible(gc()) #free up memory and report the memory usage.

```

## CALCULATE BETA DIVERSITY {.tabset .tabset-fade .tabset-pills}

I will calculate for both weighted UniFrac distances and Bray-Curtis dissimilarity index. Both indeces normalises pairwise distances/dissimilarities to be between 0 and 1, which means that while the distance between two samples will always be the same, the numerical value will depend on all the samples analysed together. Other metrics/indeces can be used, and might be relevant, but just keep in mind how the specific indeces are relevant for the interpretation of the following results.

### CREATE RAREFIED PHYLOSEQ OBJECT

Most beta diversity metrics that are based on presence/absence of bacteria are sensitive to differences in sequencing depth. The way to minimize such bias is to perform rarefaction of the data, but as rarefaction is random this introduces yet another bias. Here we will rarefy each sample multiple times and choose the most central as a representative rarefication.

```{r calc_multiple_rarefy, eval=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
load("scripts/mrarefy.Rdata")

# Load phyloseq
load(params$input)

# Perform multiple rarefactions
phy.rare <- multiple_rarefy(phy)

# Remove empty taxa
phy.rare <- prune_taxa(taxa_sums(phy.rare) > 0, phy.rare)

# Root tree
phy_tree(phy) <- ape::root(phy_tree(phy), sample(taxa_names(phy), 1), resolve.root = TRUE)

# Save object
save(phy.rare, file = "R_objects/Phyloseq_rarefied.Rdata")

# clear the environment and release memory
rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
invisible(gc()) #free up memory and report the memory usage.
```

### UNWEIGHTED UNIFRAC

The unique fraction metric, or UniFrac, measures the phylogenetic distance between sets of taxa in a phylogenetic tree as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both [(Lozupone & Knight, 2005)](https://doi.org/10.1128/AEM.71.12.8228-8235.2005). 
This metric is sensitive to sequencing depth, so it is required to use a rarefied phyloseq object
The UniFrac algorithm requires a rooted tree, so if ASVs has been removed from the raw da the tree should be rerooted manually, else a random ASV will be chosen as root.

```{r calc_UniF, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("scripts/mrarefy.Rdata")

# Root tree if necessary
if (!is.rooted(phy_tree(phy.rare))) phy_tree(phy.rare) <- ape::root(phy_tree(phy.rare), sample(taxa_names(phy.rare), 1), resolve.root = TRUE)

# Calculate UniFrac distances
unif.dist <- UniFrac(phy.rare, weighted = FALSE, parallel = FALSE)

# Calculate PCoA data
unif.pcoa <- ordinate(phy.rare, method = "PCoA",distance = unif.dist)
unif.nmds <- metaMDS(unif.dist, k = 5, trymax = 1000)

# Save distance objects
save(unif.dist, unif.nmds, unif.pcoa, file = "R_objects/UniF.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

#### SUBSETS
##### Feces
```{r calc_UniF_subs_F0248, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d2","d4","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f0248.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F0, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f0.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F2, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f2.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F4, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f4.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F08, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f08.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Feces, Cecum, Ileum
```{r calc_UniF_subs_FCEIL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.fceil8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
```{r calc_UniF_subs_CE8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Cecum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.ce8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_IL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Ileum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.il8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```



##### Feces non-Vancomycin and Vancomycin

```{r calc_UniF_subs_F0_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f0_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F0_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f0_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_UniF_subs_F2_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f2_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F2_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f2_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_UniF_subs_F4_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f4_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F4_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f4_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_UniF_subs_F8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_F8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.f8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Cecum & Ileum non-Vancomycin and Vancomycin
```{r calc_UniF_subs_CE8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.ce8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_CE8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.ce8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_IL8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.il8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_IL8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.il8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```



##### Feces diversity over time

```{r calc_UniF_subs_Feces_d0248_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "CTRL") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.feces_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_Feces_d0248_pfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.feces_pfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_Feces_d0248_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.feces_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_UniF_subs_Feces_d0248_vanpfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

tmp.dist <- as.matrix(unif.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN+PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
unif.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
unif.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = unif.sub.dist)
unif.sub.nmds <- metaMDS(unif.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, unif.sub.dist, unif.sub.nmds, unif.sub.pcoa, file = "R_objects/unif.sub.feces_vanpfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```




### WEIGHTED UNIFRAC

The unique fraction metric, or UniFrac, measures the phylogenetic distance between sets of taxa in a phylogenetic tree as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both [(Lozupone & Knight, 2005)](https://doi.org/10.1128/AEM.71.12.8228-8235.2005). 
Weighted UniFrac takes the abundance of each ASV into account instead of just presence/absence, which means that it will not be sensitive to sequencing depth. The UniFrac algorithm requires a rooted tree, so if ASVs has been removed from the raw da the tree should be rerooted manually, else a random ASV will be chosen as root.

```{r calc_wUniF, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load(params$input)

# Root tree if necessary
if (!is.rooted(phy_tree(phy))) phy_tree(phy) <- ape::root(phy_tree(phy), sample(taxa_names(phy), 1), resolve.root = TRUE)

# Calculate UniFrac distances
wuf.dist <- UniFrac(phy, weighted = TRUE, parallel = FALSE)

# Calculate PCoA data
wuf.pcoa <- ordinate(phy, method = "PCoA",distance = wuf.dist)
wuf.nmds <- metaMDS(wuf.dist, k = 5, trymax = 1000)

# Save distance objects
save(wuf.dist, wuf.nmds, wuf.pcoa, file = "R_objects/wuf.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

#### SUBSETS
##### Feces
```{r calc_wUniF_subs_F0248, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d2","d4","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f0248.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F0, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f0.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F2, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f2.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F4, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f4.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F08, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f08.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Feces, Cecum, Ileum
```{r calc_wUniF_subs_FCEIL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.fceil8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_CE8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Cecum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.ce8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_il8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Ileum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.il8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


##### Feces non-Vancomycin and Vancomycin

```{r calc_wUniF_subs_F0_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f0_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F0_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f0_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_wUniF_subs_F2_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f2_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F2_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f2_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_wUniF_subs_F4_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f4_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F4_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f4_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_wUniF_subs_F8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_F8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.f8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Cecum & Ileum non-Vancomycin and Vancomycin
```{r calc_wUniF_subs_CE8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.ce8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_CE8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.ce8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_IL8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.il8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_IL8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.il8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```



##### Feces diversity over time

```{r calc_wUniF_subs_Feces_d0248_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "CTRL") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.feces_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_Feces_d0248_pfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.feces_pfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_Feces_d0248_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.feces_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_wUniF_subs_Feces_d0248_vanpfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/wuf.RData")

tmp.dist <- as.matrix(wuf.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN+PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
wuf.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUniFrac distances

# Calculate PCoA data
wuf.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = wuf.sub.dist)
wuf.sub.nmds <- metaMDS(wuf.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa, file = "R_objects/wuf.sub.feces_vanpfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```







### BRAY-CURTIS

Bray-Curtis dissimilarity index (as implemented by the vegan package) is the sum of abundance difference for each species/ASV, divided by theoretical maximum difference between the samples if no ASV overlapped. The formula used is: $$d_{jk} = \frac{\sum|n_{ij}-n_{ik}|}{\sum(n_{ij}+n_{ik})}$$ Bray-Curtis dissimilarity is not a true distance metric as it does not adhere to the [triangle inequality](), but is often used to compare microbiomes. Bray-Curtis dissimilarities are based on the assumption that measurements are taken from equal areas, so differences in total counts between samples will bias the metric. As differences in sequences depth is due to differences in the lab procedures and not biological differences, we should transform our counts to relative abundances before calculating Bray-Curtis dissimilarities. By transforming the data to abundances no data is lost, but rarefied data can also be used.

```{r calc_bray, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load(params$input)

# transform counts to relative abundance
phy.ra <- transform_sample_counts(phy, function(x) x/sum(x))

# Calculate Bray-Curtis dissimilarities
bray.dist <- distance(phy.ra, method = "bray",)

# Calculate PCoA data
bray.pcoa <- ordinate(phy, method = "PCoA",distance = bray.dist)
bray.nmds <- metaMDS(bray.dist, k = 5, trymax = 1000)

# Save distance objects
save(bray.dist, bray.nmds, bray.pcoa, file = "R_objects/Bray.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

#### SUBSETS
##### Feces
```{r calc_Bray_subs_F0248, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d2","d4","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f0248.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F0, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f0.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F2, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f2.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F4, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f4.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F08, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f08.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Feces, Cecum, Ileum
```{r calc_Bray_subs_FCEIL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.fceil8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_CE8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Cecum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.ce8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_IL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Ileum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.il8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


##### Feces non-Vancomycin and Vancomycin

```{r calc_Bray_subs_F0_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f0_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F0_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f0_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_Bray_subs_F2_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f2_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F2_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f2_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_Bray_subs_F4_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f4_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F4_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f4_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_Bray_subs_F8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_F8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.f8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Cecum & Ileum non-Vancomycin and Vancomycin
```{r calc_Bray_subs_CE8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.ce8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_CE8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.ce8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_IL8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.il8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_IL8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.il8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```



##### Feces diversity over time

```{r calc_Bray_subs_Feces_d0248_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "CTRL") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.feces_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_Feces_d0248_pfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.feces_pfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_Feces_d0248_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.feces_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Bray_subs_Feces_d0248_vanpfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/Bray.RData")

tmp.dist <- as.matrix(bray.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN+PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
bray.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Brayrac distances

# Calculate PCoA data
bray.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = bray.sub.dist)
bray.sub.nmds <- metaMDS(bray.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, bray.sub.dist, bray.sub.nmds, bray.sub.pcoa, file = "R_objects/bray.sub.feces_vanpfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

### JACCARD

``` {r }
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load(params$input)

# transform counts
phy.ra <- transform_sample_counts(phy, function(x) x/sum(x))

# Calculate Jaccard distances
jac.dist <- distance(phy.ra, method = "jaccard", binary = TRUE)

# Calculate PCoA data
jac.pcoa <- ordinate(phy, method = "PCoA",distance = jac.dist)
jac.nmds <- metaMDS(jac.dist, k = 5, trymax = 1000)

# Save distance objects
save(jac.dist, jac.nmds, jac.pcoa, file = "R_objects/jac.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())
```

#### SUBSETS
##### Feces
```{r calc_Jac_subs_F0248, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d2","d4","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate UniFrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f0248.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F0, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f0.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F2, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f2.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F4, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f4.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F08, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day %in% c("d0","d8"))

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f08.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Feces, Cecum, Ileum
```{r calc_Jac_subs_FCEIL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.fceil8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_CE8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Cecum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.ce8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_IL8, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, day == "d8" & material == "Ileum")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate wUnifrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.il8.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


##### Feces non-Vancomycin and Vancomycin

```{r calc_Jac_subs_F0_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f0_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F0_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d0" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f0_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_Jac_subs_F2_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f2_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F2_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d2" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f2_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_Jac_subs_F4_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f4_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F4_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d4" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f4_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


```{r calc_Jac_subs_F8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_F8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.f8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```
##### Cecum & Ileum non-Vancomycin and Vancomycin
```{r calc_Jac_subs_CE8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.ce8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_CE8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Cecum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.ce8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_IL8_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "ctrl")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.il8_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_IL8_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Ileum" & day == "d8" & van == "van")

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.il8_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```



##### Feces diversity over time

```{r calc_Jac_subs_Feces_d0248_ctrl, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "CTRL") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.feces_ctrl.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_Feces_d0248_pfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.feces_pfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_Feces_d0248_van, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.feces_van.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```

```{r calc_Jac_subs_Feces_d0248_vanpfos, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/jac.RData")

tmp.dist <- as.matrix(jac.dist)
# Create subset
phy.sub <- subset_samples(phy.rare, material == "Feces" & treatment == "VAN+PFOS") 

SUBS <- sample_names(phy.sub)

tmp.sub.dist <- tmp.dist[SUBS, SUBS]
jac.sub.dist <- as.dist(tmp.sub.dist,diag = TRUE, upper = TRUE)

# Calculate Jacrac distances

# Calculate PCoA data
jac.sub.pcoa <- ordinate(phy.sub, method = "PCoA",distance = jac.sub.dist)
jac.sub.nmds <- metaMDS(jac.sub.dist, k = 5, trymax = 1000)

# Save distance objects
save(phy.sub, jac.sub.dist, jac.sub.nmds, jac.sub.pcoa, file = "R_objects/jac.sub.feces_vanpfos.RData")

# clear the environment and release memory
rm(list = ls(all.names = TRUE))
invisible(gc())

```


## BATCH EFFECTS {.tabset .tabset-fade .tabset-pills}

For the beta diversity we will also have to test for batch effects, and if they are significant, correct for them when performing the project relevant analyses.
I will test each diversity index individually, so if additional indeces have been included, just copy and and adapt the following sections. Additional tests only including feces samples.

### UNIFRAC

```{r batch_UniF, eval=TRUE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load("R_objects/Phyloseq_rarefied.Rdata")
load("R_objects/UniF.RData")

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy.rare))

# Run PERMANOVA for batch variable
FORMULA <- as.formula(paste("unif.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(unif.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# Subset for Feces
sub.phy <- subset_samples(phy.rare, material == "Feces")
mdat <- data.frame(sample_data(sub.phy))
load("R_objects/unif.sub.f0248.RData")

# Run PERMANOVA for batches only with Feces samples
FORMULA <- as.formula(paste("unif.sub.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(unif.sub.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
Batch effect is observed between c2 and other two batches, as can be expected from difference in material origin (feces vs. cecum and ileum). Significant batch effect is observed between c3-c4 in unweighted UniFrac which should be taken into account in analysis between fecal samples over days d0/d8 to d2/d4. Due to the distribution of samples over batches, a difference is  expected as the number of samples reflecting vancomycin treatment were different between the batches (higher portion of samples not exposed to vancomycin in c4).


### WEIGHTED UNIFRAC

```{r batch_WUF, eval=TRUE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load(params$input)
load("R_objects/wuf.RData")

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy))

# Run PERMANOVA for batch variable
FORMULA <- as.formula(paste("wuf.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(wuf.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# Subset for Feces
sub.phy <- subset_samples(phy, material == "Feces")
mdat <- data.frame(sample_data(sub.phy))
load("R_objects/wuf.sub.f0248.RData")

# Run PERMANOVA for batches only with Feces samples
FORMULA <- as.formula(paste("wuf.sub.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(wuf.sub.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
Batch effect is observed between c2 and other two batches, as can be expected from difference in material origin (feces vs. cecum and ileum). No batch effect is found between c3-c4 for weighted UniFrac.

### BRAY-CURTIS

```{r batch_Bray, eval=TRUE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load(params$input)
load("R_objects/Bray.RData")

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy))

# Run PERMANOVA for batch variable
FORMULA <- as.formula(paste("bray.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(bray.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# Subset for Feces
sub.phy <- subset_samples(phy, material == "Feces")
mdat <- data.frame(sample_data(sub.phy))
load("R_objects/bray.sub.f0248.RData")

# Run PERMANOVA for batches only with Feces samples
FORMULA <- as.formula(paste("bray.sub.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(bray.sub.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
Batch effect is observed between c2 and other two batches, as can be expected from difference in material origin (feces vs. cecum and ileum). No batch effect is found between c3-c4 for Bray-Curtis.

### JACCARD

```{r batch_Jac, eval=TRUE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# load
load(params$input)
load("R_objects/jac.RData")

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy))

# Run PERMANOVA for batch variable
FORMULA <- as.formula(paste("jac.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(jac.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# Subset for Feces
sub.phy <- subset_samples(phy, material == "Feces")
mdat <- data.frame(sample_data(sub.phy))
load("R_objects/jac.sub.f0248.RData")

# Run PERMANOVA for batches only with Feces samples
FORMULA <- as.formula(paste("jac.sub.dist ~ ", params$batch, sep = ""))
batch.PERM <- adonis2(FORMULA, data = mdat)

# Compare the betadiversity dispertion for the batch variable
batch.bdisp <- betadisper(jac.sub.dist, mdat[,params$batch])
anova(batch.bdisp)

# If significant a post hoc test can compare pairwise
TukeyHSD(batch.bdisp)

plot(batch.bdisp)
boxplot(batch.bdisp)

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
Batch effect is observed between c2 and other two batches, as can be expected from difference in material origin (feces vs. cecum and ileum). Significant batch effect is observed between c3-c4 in Jaccard which should be taken into account in analysis between fecal samples over days d0/d8 to d2/d4. Due to the distribution of samples over batches, a difference is expected as the number of samples reflecting vancomycin treatment were different between the batches (higher portion of samples not exposed to vancomycin in c4).

### OVERALL INTERPRETATION

There were significant batch dependent differences in beta diversity in all investigated metrices, which need to be considered for further analysis. Due to sequencial run of samples batches the sample distribution for samples are not uniformly randomized on all three batches, rather the "c2" contains only samples from ileum and cecum luminal content, while "c3" and "c4" contains fecal samples of day 2-4 and day 0+8, respectively. This should be taken into account when comparing data between day 0 or 8 with day 2 or 4.  Also, comparison with ileal and cecal samples should be weighted against this. Due to the distribution of samples over batches, a difference is expected as the number of samples reflecting vancomycin treatment were different between the batches c3 and c4 (higher portion of samples not exposed to vancomycin in c4).


# GROUP-WISE TREATMENT - CATEGORICAL VARIABLES EFFECTS {.tabset .tabset-fade .tabset-pills}

In this first section (MANUAL) individual variables and metrices can be tested and plotted manually.
In the second following section (PCOA ITERATIONS) all calculated iterations created during subsetting are analysed to create principal coordinate analyses through a for-loop created to select for appropriate variables and color-schemes for each iteration.

## MANUAL
This first part loads the data and defines beta-diversity metric and tested variable for the following. Function ggExtra::ggMarginal is used for marginal boxplots for axis 1 and axis 2.
**OBS** NMDS has been removed/silenced in the script below

```{r , eval=TRUE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")

# Choose metric
METRIC <- "Bray"     # Metric for analysis: "Bray", "UniF", "WUF", "Jaccard"
# Choose variable 
VAR <- "treatment"   # Categorical factor - main
VAR2 <- "day"        # Categorical factor 2
DAY <- "d8"          # Values for "day": "d0", "d2", "d4", "d8"
MTRL <- "Feces"      # Materials: "Feces", "Cecum", "Ileum"
VAN <- "van"         # Variable for without and with vancomycin treatment: ctrl and van. Should be removed for test on all four treatment groups.
SUBNAME <- "f8_van"  # Subname should match the filename format for iterations used above: { METRIC.sub.SUBNAME.RData }
# Color scheme
COL <-  c("CTRL" = "#606060","PFOS" = "#31b44b","VAN" = "#6699d1","VAN+PFOS" = "#efc000")

# Create subfolder if needed
sub.dir <- paste0("plots/bdiv_manual/beta_",SUBNAME)
if (!file.exists(sub.dir)) dir.create(file.path(getwd(), sub.dir))
sdir <- paste0(sub.dir,"/")

# Load data based on METRIC
load(params$input)

if (METRIC == "UniF") {
  filename <- paste0("R_objects/unif.sub.",SUBNAME,".RData")
  load(filename)
  dist.used <- unif.sub.dist
  nmds.used <- unif.sub.nmds
  pcoa.used <- unif.sub.pcoa
  rm(unif.sub.dist, unif.sub.nmds, unif.sub.pcoa)
} else if (METRIC == "WUF") {
  filename <- paste0("R_objects/wuf.sub.",SUBNAME,".RData")
  load(filename)
  dist.used <- wuf.sub.dist
  nmds.used <- wuf.sub.nmds
  pcoa.used <- wuf.sub.pcoa
  rm(wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa)
} else if (METRIC == "Bray"){
  filename <- paste0("R_objects/bray.sub.",SUBNAME,".RData")
  load(filename)
  dist.used <- bray.sub.dist
  nmds.used <- bray.sub.nmds
  pcoa.used <- bray.sub.pcoa
  rm(bray.sub.dist, bray.sub.nmds, bray.sub.pcoa)
} else if (METRIC == "Jaccard"){
  filename <- paste0("R_objects/jac.sub.",SUBNAME,".RData")
  load(filename)
  dist.used <- jac.sub.dist
  nmds.used <- jac.sub.nmds
  pcoa.used <- jac.sub.pcoa
  rm(jac.sub.dist, jac.sub.nmds, jac.sub.pcoa)
}

# Extract metadata from phyloseq
mdat <- data.frame(sample_data(phy.sub))

# Subset the metadata based on the above
mdat <- mdat[mdat$day == DAY & mdat$material == MTRL,]

# If a variable consist of numbers, but represent distinct groups remember to make it into a factor
mdat[,VAR] <- as.factor(mdat[,VAR])

## STATISTICAL TEST
# Compare the betadiversity dispertion for Weighted UniFrac
bdisp <- betadisper(dist.used, mdat[,VAR], bias.adjust=TRUE)
test.anova <- anova(bdisp)
test.anova

# dispertion by group
b.title <- paste0(METRIC," ",SUBNAME)
p.bdisp <- boxplot(bdisp, main = b.title)


# Test which groups differ (only if the anova test was significant)
(HSD <- TukeyHSD(bdisp))
p.HSD <- plot(HSD)

# Run PERMANOVA for the variable
FORMULA <- as.formula(paste("dist.used ~", VAR, sep = " "))
(perm.test <- adonis2(FORMULA, data = mdat, permutations = 9999,na.action = na.omit))
FORMULA2 <- as.formula(paste("dist.used ~", "pfos", sep = " "))
(perm.test2 <- adonis2(FORMULA2, data = mdat, permutations = 9999,na.action = na.omit))
# Use vegan to test how well metadata fits ordination
fit.out <- envfit(nmds.used, mdat[,c(VAR,VAR2)],na.rm=TRUE)
fit.out

# ### EIGENVALUES
# For PCoA each axis represent a specific amount of the overall variation in the dataset. this information can easily be extracted and plotted.

# Extract eigen values
eigen <- pcoa.used$values
eigen$Axis <- as.numeric(row.names(eigen))

# Create plots for both distance indeces
p.eigen <- ggplot(eigen[1:10,], aes(x = as.factor(Axis), y = 100*Rel_corr_eig)) + 
  geom_col(aes(fill = as.factor(Axis))) +
  geom_point(aes(x = Axis, y = 100*Cum_corr_eig)) +
  geom_line(aes(x = Axis, y = 100*Cum_corr_eig)) +
  ylab("Variance explained (%)") +
  xlab("Axis") +
  theme_pubr(legend = "none") + ggsci::scale_fill_jco()
p.eigen + ggtitle(paste("Variance per axis for", METRIC, sep = " "))
suppressMessages(ggsave(plot = p.eigen, filename = paste0(sdir,"bdiv_eigen_",METRIC,"_",SUBNAME,".png"), device = "png"))
suppressMessages(ggsave(plot = p.eigen, filename = paste0(sdir,"bdiv_eigen_",METRIC,"_",SUBNAME,".pdf"), device = "pdf"))

### ORDINATION

# Phyloseq has a plotting function, but it is a bit limited in some of the settings. Therefore, I recommend to use the function to create a table of the data and then make your own plots
# The first plot highlights the location of each group on the first 5 axis. Based on this an optimal set of axis can be chosen for the following ordination plot.

# Create plots of eigenvalues for PCoA plots
pcoa.tab <- plot_ordination(phy, pcoa.used,axes = 1:5,justDF = TRUE)
nmds.tab <- plot_ordination(phy, nmds.used,axes = 1:5,justDF = TRUE)

# Reformat tables to create one common table
colnames(nmds.tab)[1:5] <- c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5")

nmds.tab$ordination <- "nmds"
pcoa.tab$ordination <- "pcoa"

ord.tab <- rbind(nmds.tab,pcoa.tab)
ord.tab$treatment <- as.factor(ord.tab$treatment)

# Melt axis to be in one variable
axis.tab <- pivot_longer(data = ord.tab, cols = c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5"), names_to = "Axis", values_to = "position")

# Plot positions on axes
plot.pos <- ggplot(axis.tab, aes_string(x = "ordination", y = "position", fill = VAR)) +
  geom_boxplot() +
  facet_grid(Axis~.) +
  coord_flip() + 
  theme_pubr(legend = "bottom") + ggsci::scale_fill_jco()
TITLE <- paste0(METRIC," - Axis effect: ",MTRL," ","day ",DAY)
plot.pos2 <- print(plot.pos + ggtitle(TITLE))
plot.pos2
suppressMessages(ggsave(plot = plot.pos2, filename = paste0(sdir,"bdiv_axis_effect_",METRIC,"_",SUBNAME,".png"), device = "png"))
suppressMessages(ggsave(plot = plot.pos2, filename = paste0(sdir,"bdiv_axis_effect_",METRIC,"_",SUBNAME,".pdf"), device = "pdf"))

# Create PCoA plot
p <- ggplot(ord.tab[ord.tab$ordination == "pcoa",], aes_string(x = "Axis.1", y = "Axis.2", color = VAR, fill = VAR, group = VAR))
p <- p + geom_vline(xintercept = c(0), color = "grey70", alpha = 0.6, linetype = 1, linewidth = 0.5) +
  geom_hline(yintercept = c(0), color = "grey70", alpha = 0.6, linetype = 1, linewidth = 0.5) +
  geom_point() +
  theme_pubr(legend = "bottom") +
  labs(color=VAR) +
  stat_ellipse(geom = "polygon", alpha = 0.1) +
  scale_color_manual(values = COL) +
  scale_fill_manual(values = COL) +
  border(color = "black",size = 0.5) +
  labs(color = "Treatment") +
  guides(fill = "none")
p.pcoa <- ggExtra::ggMarginal(p = p, type = 'boxplot', size = 10, groupFill = TRUE)
p.pcoa

# Save plot
plotfile <- file.path("plots/bdiv",paste0("bdiv_ordination_",VAR,".png"))
ggsave(filename = plotfile, plot = p.pcoa, width = 150, height = 150, device = "png", units = "mm",dpi = 300)
# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

## ALL PCOA ITERATIONS {.tabset .tabset-fade .tabset-pills}
For-loop setup to generate analysis and PCoA visualization for all listed iterations. Eigenvalue plot is excluded from the loop due to constant errors related to calculations of "Rel_corr_eig". Function ggExtra::ggMarginal is used for marginal boxplots for axis 1 and axis 2.
**OBS** NMDS has been removed/silenced in the script below

```{r , eval=TRUE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")
# Default variables
VAR <- "treatment"
VAR2 <- "day"
# Default color scheme for treatment groups 
COL <-  c("CTRL" = "#606060","PFOS" = "#31b44b","VAN" = "#6699d1","VAN+PFOS" = "#efc000")

# Create string of iterations based on subsets created above = SUBNAME
str.subnameT<- c("feces_ctrl","feces_pfos","feces_van","feces_vanpfos",
                 "f0","f0_ctrl","f0_van","f2","f2_ctrl","f2_van",
                 "f4","f4_ctrl","f4_van",
                 "f8","f8_ctrl","f8_van",
                 "il8","il8_ctrl","il8_van",
                 "ce8","ce8_ctrl","ce8_van",
                 "f08","f0248")

# Determine variables for sub-setting and naming the run
for (SUBNAME in str.subnameT) {
  VAR <- "treatment"
  VAR2 <- "day"
  COL <-  c("CTRL" = "#606060","PFOS" = "#31b44b","VAN" = "#6699d1","VAN+PFOS" = "#efc000")
  print(paste0("Processing ",SUBNAME,"..."))
  if (grepl("feces_",SUBNAME) == TRUE) {
    MTRL <- "Feces"
    MN <- MTRL
    VAR <- "day"
    VAR2 <- "treatment"
    DAY <- c("d0","d2","d4","d8")
    dn <- "All days"
  } else if (grepl("fceil",SUBNAME) == TRUE) {
    MTRL <- c("Feces","Cecum","Ileum")
    MN <- "All materials"
    VAR <- "material"
    VAR2 <- "treatment"
    COL <- c("Feces" = "#cd534c","Cecum" = "#efc000","Ileum" = "#0073c2")
  } else if (grepl("f",SUBNAME) == TRUE) {
    MTRL <- "Feces"
    MN <- MTRL
  } else if (grepl("ce",SUBNAME) == TRUE) {
    MTRL <- "Cecum"
    MN <- MTRL
  } else if (grepl("il",SUBNAME) == TRUE) {
    MTRL <- "Ileum"
    MN <- MTRL
  }
  if (grepl("feces_",SUBNAME) == TRUE) {
    if (grepl("_ctrl",SUBNAME) == TRUE) {
      COL <- c("d0" = "#a0a0a0","d2" = "#808080","d4" = "#606060","d8" = "#494949")
      print("Treatment sorting. No vancomycin definition needed")
    } else if (grepl("_pfos",SUBNAME) == TRUE) {
      COL <- c("d0" = "#48ff62","d2" = "#45ce4b","d4" = "#31b44b","d8" = "#278937")
      print("Treatment sorting. No vancomycin definition needed")
    } else if (grepl("_vanpfos",SUBNAME) == TRUE) {
      COL <- c("d0" = "#f7e789","d2" = "#f7e054","d4" = "#fcdd05","d8" = "#c4a110")
      print("Treatment sorting. No vancomycin definition needed")
    } else if (grepl("_van",SUBNAME) == TRUE) {
      COL <- c("d0" = "#8bd2fc","d2" = "#81c3ea","d4" = "#6699d1","d8" = "#487baa")
      print("Treatment sorting. No vancomycin definition needed")
    }
  } else if (grepl("ctrl", SUBNAME) == TRUE) {
    STA <- "ctrl"
  } else if (grepl("van", SUBNAME) == TRUE) {
    STA <- "van"
  } else {
  }
  if (grepl("0248",SUBNAME) == TRUE) {
    DAY <- c("d0","d2","d4","d8")
    dn <- "All days"
  } else if (grepl("08",SUBNAME) == TRUE) {
    DAY <- c("d0","d8")
    dn <- "Day 0 + 8"
  } else if (grepl("0", SUBNAME) == TRUE) {
    DAY <- "d0"
    dn <- "Day 0"
  } else if (grepl("2", SUBNAME) == TRUE) {
    DAY <- "d2"
    dn <- "Day 2"
  } else if (grepl("4", SUBNAME) == TRUE) {
    DAY <- "d4"
    dn <- "Day 4"
  } else if (grepl("8", SUBNAME) == TRUE) {
    DAY <- "d8"
    dn <- "Day 8"
  }

  sub.dir <- paste0("plots/bdiv/beta_",SUBNAME)
  if (!file.exists(sub.dir)) dir.create(file.path(getwd(), sub.dir))
  sdir <- paste0(sub.dir,"/")

  for (METRIC in params$metrices) {
  # Load data
    print(paste0("Processing ",SUBNAME," in ",METRIC,"..."))
    load(params$input)
    #METRIC <- "UniF"
    if (METRIC == "UniF") {
      filename <- paste0("R_objects/unif.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- unif.sub.dist
      nmds.used <- unif.sub.nmds
      pcoa.used <- unif.sub.pcoa
      rm(unif.sub.dist, unif.sub.nmds, unif.sub.pcoa)
    } else if (METRIC == "WUF") {
      filename <- paste0("R_objects/wuf.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- wuf.sub.dist
      nmds.used <- wuf.sub.nmds
      pcoa.used <- wuf.sub.pcoa
      rm(wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa)
    } else if (METRIC == "Bray"){
      filename <- paste0("R_objects/bray.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- bray.sub.dist
      nmds.used <- bray.sub.nmds
      pcoa.used <- bray.sub.pcoa
      rm(bray.sub.dist, bray.sub.nmds, bray.sub.pcoa)
    } else if (METRIC == "Jaccard"){
      filename <- paste0("R_objects/jac.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- jac.sub.dist
      nmds.used <- jac.sub.nmds
      pcoa.used <- jac.sub.pcoa
      rm(jac.sub.dist, jac.sub.nmds, jac.sub.pcoa)
    }
    
    # Extract metadata from phyloseq
    mdat <- data.frame(sample_data(phy.sub))
    
    # Subset the metadata based on the above
    if (grepl("feces_",SUBNAME) == TRUE) {
      mdat <- mdat[mdat$day %in% DAY & mdat$material %in% MTRL,]
    } else if (grepl("_",SUBNAME) == FALSE) {
      mdat <- mdat[mdat$day %in% DAY & mdat$material %in% MTRL,]
    } else {
      mdat <- mdat[mdat$day %in% DAY & mdat$material %in% MTRL & mdat$van %in% STA,]
    }
    
    # If a variable consist of numbers, but represent distinct groups remember to make it into a factor
    mdat[,VAR] <- as.factor(mdat[,VAR])
    
    ## STATISTICAL TEST
    # Compare the betadiversity dispertion for Weighted UniFrac
    bdisp <- betadisper(dist.used, mdat[,VAR], bias.adjust=TRUE)
    test.anova <- anova(bdisp)
    test.anova
    
    # dispertion by group
    b.title <- paste0(METRIC," ",SUBNAME)
    p.bdisp <- boxplot(bdisp, main = b.title)
    
    png(file = paste0(sdir,"bdiv_bdisp_boxplot_",METRIC,"_",SUBNAME,".png"))
    boxplot(bdisp, main = b.title)
    dev.off()
    
    # Test which groups differ (only if the anova test was significant)
    (HSD <- TukeyHSD(bdisp))
    p.HSD <- plot(HSD)
    
    png(file = paste0(sdir,"bdiv_bdisp_HSD_",METRIC,"_",SUBNAME,".png"))
    plot(HSD)
    dev.off()
    
    # Run PERMANOVA for the variable
    FORMULA <- as.formula(paste("dist.used ~", VAR, sep = " "))
    (perm.test <- adonis2(FORMULA, data = mdat, permutations = 9999,na.action = na.omit))
    
    # Use vegan to test how well metadata fits ordination
    fit.out <- envfit(nmds.used, mdat[,c(VAR,VAR2)],na.rm=TRUE)
    fit.out
    
    stat.out <- paste0(sdir,"bdiv_bdisp_",METRIC,"_",SUBNAME,".txt")
    sink(stat.out)
    print(test.anova)
    print(perm.test)
    print(fit.out)
    sink()
    
    ### ORDINATION
    # Create plots of eigenvalues for PCoA plots
    pcoa.tab <- plot_ordination(phy, pcoa.used,axes = 1:5,justDF = TRUE)
    #nmds.tab <- plot_ordination(phy, nmds.used,axes = 1:5,justDF = TRUE)
    
    # Reformat tables to create one common table
    #colnames(nmds.tab)[1:5] <- c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5")
    
    #nmds.tab$ordination <- "nmds"
    pcoa.tab$ordination <- "pcoa"
    
    ord.tab <- rbind(pcoa.tab) #nmds.tab,
    ord.tab$treatment <- as.factor(ord.tab$treatment)
    
    # Melt axis to be in one variable
    axis.tab <- pivot_longer(data = ord.tab, cols = c("Axis.1","Axis.2","Axis.3","Axis.4","Axis.5"), names_to = "Axis", values_to = "position")
    
    # Plot positions on axes
    plot.pos <- ggplot(axis.tab, aes_string(x = "ordination", y = "position", fill = VAR)) +
      geom_boxplot() +
      facet_grid(Axis~.) +
      coord_flip() + 
      theme_pubr(legend = "bottom") + scale_fill_manual(values = COL)#ggsci::scale_fill_jco()
    plot.posT <- plot.pos + ggtitle(paste0(METRIC," - Axis effect: ",MN," ",dn))
    suppressMessages(ggsave(plot = plot.posT, filename = paste0(sdir,"bdiv_axis_effect_",METRIC,"_",SUBNAME,".png"), device = "png", units = "mm", width = 200, height = 100, dpi = 300))
    suppressMessages(ggsave(plot = plot.posT, filename = paste0(sdir,"bdiv_axis_effect_",METRIC,"_",SUBNAME,".pdf"), device = "pdf", units = "mm", width = 200, height = 100, dpi = 300))
    
    # Create ordination plots
    p <- ggplot(ord.tab[ord.tab$ordination == "pcoa",], aes_string(x = "Axis.1", y = "Axis.2", color = VAR, fill = VAR, group = VAR))
    p <- p + 
      #geom_vline(xintercept = c(0), color = "grey70", alpha = 0.4, linetype = 1, linewidth = 0.5) +
      #geom_hline(yintercept = c(0), color = "grey70", alpha = 0.4, linetype = 1, linewidth = 0.5) +
      geom_point() +
      theme_pubr(legend = "bottom") +
      labs(color=VAR) +
      stat_ellipse(geom = "polygon", alpha = 0.1) +
      scale_color_manual(values = COL) +
      scale_fill_manual(values = COL) +
      border(color = "black",size = 0.5) +
      labs(color = "Treatment") +
      guides(fill = "none")
    plot.beta <- ggExtra::ggMarginal(p = p, type = 'boxplot', size = 10, groupFill = TRUE)
      
    suppressMessages(ggsave(filename = paste0(sdir,"bdiv_PCoA_",METRIC,"_",SUBNAME,".png"), plot.beta, device = "png", units = "mm", width = 150, height = 150, dpi = 300))
    suppressMessages(ggsave(filename = paste0(sdir,"bdiv_PCoA_",METRIC,"_",SUBNAME,".pdf"), plot.beta, device = "pdf", units = "mm", width = 150, height = 150, dpi = 300))

  message(paste0("Finished processing ",SUBNAME," without issues. Moving on..."))
  } 
  # clear the environment and release memory except variables used for next run
  rm(list = ls(all.names = TRUE)[!ls(all.names = TRUE) %in% c("params","VAR","VAR2","str.subnameT","COL")])
  invisible(gc())
}
# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())
```

# PERMANOVA
In this section the contribution of categorical values for treatment ("pfos" and "van" = vancomycin) to beta diversity distances for each metric is tested using PERMANOVA test (vegan::adonis2). Tests for each sampling day are tested and for subsets based on vancomycin treatment.

## PERMANOVA FECES DAY 0
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 0"

# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f0"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d0",]
perm.unif <- adonis2(unif.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f0"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d0",]
perm.wuf <- adonis2(wuf.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f0"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d0",]
perm.bray <- adonis2(bray.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f0"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d0",]
perm.jac <- adonis2(jac.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.jac

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA FECES DAY 2
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 2"

# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f2"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2",]
perm.unif <- adonis2(unif.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f2"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2",]
perm.wuf <- adonis2(wuf.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f2"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2",]
perm.bray <- adonis2(bray.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f2"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2",]
perm.jac <- adonis2(jac.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.jac

as.data.frame(perm.unif)["pfos","Pr(>F)"]
coef <- coefficients(perm.unif)["",]


# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA FECES day 2 CTRL & VAN
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 2"
VAN <- "no vancomycin"

# No vancomycin treatment
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f2_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "ctrl",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f2_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "ctrl",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f2_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "ctrl",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f2_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "ctrl",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac
message("")
# With vancomycin treatment
VAN <- "vancomycin"
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f2_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "van",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f2_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "van",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f2_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "van",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f2_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d2" & dat$van == "van",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA FECES DAY 4
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 4"

# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f4"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4",]
perm.unif <- adonis2(unif.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f4"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4",]
perm.wuf <- adonis2(wuf.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f4"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4",]
perm.bray <- adonis2(bray.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f4"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4",]
perm.jac <- adonis2(jac.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.jac

as.data.frame(perm.unif)["pfos","Pr(>F)"]
coef <- coefficients(perm.unif)["",]


# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA FECES day 4 CTRL & VAN
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 4"
VAN <- "no vancomycin"

# No vancomycin treatment
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f4_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "ctrl",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f4_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "ctrl",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f4_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "ctrl",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f4_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "ctrl",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac
message("")
# With vancomycin treatment
VAN <- "vancomycin"
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f4_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "van",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f4_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "van",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f4_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "van",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f4_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d4" & dat$van == "van",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA FECES DAY 8
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 8"

# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8",]
perm.unif <- adonis2(unif.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8",]
perm.wuf <- adonis2(wuf.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8",]
perm.bray <- adonis2(bray.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8",]
perm.jac <- adonis2(jac.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.jac

as.data.frame(perm.unif)["pfos","Pr(>F)"]
coef <- coefficients(perm.unif)["",]


# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA FECES day 8 CTRL & VAN
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Feces"
DAY <- "Day 8"
VAN <- "no vancomycin"

# No vancomycin treatment
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "ctrl",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "ctrl",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "ctrl",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "ctrl",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac
message("")
# With vancomycin treatment
VAN <- "vancomycin"
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.f8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "van",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.f8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "van",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.f8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "van",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.f8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Feces" & dat$day == "d8" & dat$van == "van",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA CECUM DAY 8
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Cecum"
DAY <- "Day 8"

# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.ce8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8",]
perm.unif <- adonis2(unif.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.ce8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8",]
perm.wuf <- adonis2(wuf.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.ce8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8",]
perm.bray <- adonis2(bray.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.ce8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8",]
perm.jac <- adonis2(jac.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.jac

as.data.frame(perm.unif)["pfos","Pr(>F)"]
coef <- coefficients(perm.unif)["",]


# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA CECUM DAY 8 CTRL & VAN
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Cecum"
DAY <- "Day 8"
VAN <- "no vancomycin"

# No vancomycin treatment
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.ce8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "ctrl",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.ce8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "ctrl",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.ce8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "ctrl",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.ce8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "ctrl",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac
message("")
# With vancomycin treatment
VAN <- "vancomycin"
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.ce8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "van",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.ce8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "van",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.ce8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "van",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.ce8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Cecum" & dat$day == "d8" & dat$van == "van",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA ILEUM DAY 8
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Ileum"
DAY <- "Day 8"

# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.il8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8",]
perm.unif <- adonis2(unif.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.il8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8",]
perm.wuf <- adonis2(wuf.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.il8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8",]
perm.bray <- adonis2(bray.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.il8"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8",]
perm.jac <- adonis2(jac.sub.dist ~ van*pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY))
perm.jac

as.data.frame(perm.unif)["pfos","Pr(>F)"]
coef <- coefficients(perm.unif)["",]


# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```
## PERMANOVA ILEUM DAY 8 CTRL & VAN
``` {r , eval=TRUE, echo=TRUE}
MTRL <- "Ileum"
DAY <- "Day 8"
VAN <- "no vancomycin"

# No vancomycin treatment
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.il8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "ctrl",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.il8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "ctrl",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.il8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "ctrl",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.il8_ctrl"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "ctrl",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac
message("")
# With vancomycin treatment
VAN <- "vancomycin"
message(paste0("PERMANOVA distance test for PFOS effect on ",MTRL," ",DAY," with ",VAN))
# UniF
METRIC <- "UniF"
SUBNAME <- "unif.sub.il8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "van",]
perm.unif <- adonis2(unif.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.unif
message("")
# WUF
METRIC <- "WUF"
SUBNAME <- "wuf.sub.il8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "van",]
perm.wuf <- adonis2(wuf.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.wuf
message("")
# BRAY
METRIC <- "Bray"
SUBNAME <- "bray.sub.il8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "van",]
perm.bray <- adonis2(bray.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.bray
message("")
# Jaccard
METRIC <- "Jaccard"
SUBNAME <- "jac.sub.il8_van"
load(paste0("R_objects/",SUBNAME,".RData"))
load(params$input)

dat <- data.frame(sample_data(phy))
dat <- dat[dat$material == "Ileum" & dat$day == "d8" & dat$van == "van",]
perm.jac <- adonis2(jac.sub.dist ~ pfos, data = dat, permutations = 9999, na.action = na.omit)
message(paste0("PERMANOVA on ",METRIC,": ",MTRL," ",DAY," ",VAN))
perm.jac

# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

# ANALYSES NOT INCLUDED IN PUBLICATION

Following code chunks have been used in the processing of data but are not used for content of the paper.

## DISTANCE COMPARISON BOXPLOTS
Section under development for creating a statistic and visual for distance differences between treatment groups or days for each iteration created above. First section is a manual setup followed by a section with a for-loop generating plots for each iteration listed.
THIS SECTION SHOULD MOST LIKELY BE DELETED FROM FINAL SCRIPT.
``` {r beta_boxplots_bray, eval=F, echo=T}
params <- readRDS("R_objects/bdiv_params.RDS")
load("R_objects/Phyloseq_rarefied.RData")
# load(params$input)
METRIC <- "Bray"
SUBNAME <- "f8_van"
MTRL <- "Feces"
DAY <- "d8"
VAN <- "van"

COL <- c("#606060","#31b44b")

# SUBSET
phy.sub <- subset_samples(phy.rare, material == MTRL & day %in% DAY & van == VAN)

# Transform original data
phy.rl <- transform_sample_counts(phy.sub, function(x) x/sum(x))
phy.rl
head(otu_table(phy.rl)[,1:6])
bray.dist <- distance(phy.rl, method = "bray",)

# Create matrix from distances
bray.m <- melt(as.matrix(bray.dist))
bray.m

# Remove self-comparisons
bray.m <- bray.m %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)
bray.m

# Get sample data
sd <- data.frame(sample_data(phy.rl))
sd <- sd %>%
select(ID, treatment) %>%
mutate_if(is.factor,as.character)
sd

# Combine distances with sample data
colnames(sd) <- c("Var1","Treat1")
bray.sd <- left_join(bray.m, sd, by = "Var1")

colnames(sd) <- c("Var2","Treat2")
bray.sd <- left_join(bray.sd,sd,by = "Var2")

#bray.sd$group <- paste(bray.sd$Treat1, bray.sd$Treat2, sep = "_")
bray.sd

write.csv(file = paste0("output/bdiv_comparison_boxplot_",MTRL,"_",DAY,"_",VAN,"_",METRIC,"_.csv"),bray.sd)

# Create boxplot
p <- ggplot(bray.sd, aes(x = Treat2, y = value)) +
  theme_bw() +
  geom_boxplot(aes(color = ifelse(Treat1 == Treat2, "red", "black"))) +
  geom_point(position = position_jitter(height = 0.2,width = 0.2)) +
  scale_color_identity() +
  facet_wrap(~ Treat1, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle(paste0("Distance metric = ",METRIC)) +
  labs(x ="Treatment",y="Distance")
p
#p.stat <- p + stat_pvalue_manual(stat.sig, label = "p.adj.format",tip.length = 0)

suppressMessages(ggsave(filename = paste0("plots/bdiv/beta_",SUBNAME,"/",METRIC,"_boxplot_",SUBNAME,".png"), p, device = "png", units = "mm", width = 350, height = 200, dpi = 300))
# clear the environment and release memory
rm(list = ls(all.names = TRUE)[ls(all.names = TRUE) != "params"])
invisible(gc())

```

## DISTANCE COMPARISON ALL
``` {r beta_boxplots_all, eval=FALSE, echo=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")

str.subnameT<- c("feces_ctrl","feces_pfos","feces_van","feces_vanpfos",
                 "f0","f0_ctrl","f0_van","f2","f2_ctrl","f2_van",
                 "f4","f4_ctrl","f4_van",
                 "f8","f8_ctrl","f8_van",
                 "il8","il8_ctrl","il8_van",
                 "ce8","ce8_ctrl","ce8_van",
                 "f08","f0248","fceil8")

for (SUBNAME in str.subnameT) {
  for (METRIC in params$metrices) {
  # Load data
    load(params$input)
    #METRIC <- "UniF"
    if (METRIC == "UniF") {
      filename <- paste0("R_objects/unif.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- unif.sub.dist
      nmds.used <- unif.sub.nmds
      pcoa.used <- unif.sub.pcoa
      rm(unif.sub.dist, unif.sub.nmds, unif.sub.pcoa)
    } else if (METRIC == "WUF") {
      filename <- paste0("R_objects/wuf.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- wuf.sub.dist
      nmds.used <- wuf.sub.nmds
      pcoa.used <- wuf.sub.pcoa
      rm(wuf.sub.dist, wuf.sub.nmds, wuf.sub.pcoa)
    } else if (METRIC == "Bray"){
      filename <- paste0("R_objects/bray.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- bray.sub.dist
      nmds.used <- bray.sub.nmds
      pcoa.used <- bray.sub.pcoa
      rm(bray.sub.dist, bray.sub.nmds, bray.sub.pcoa)
    } else if (METRIC == "Jaccard"){
      filename <- paste0("R_objects/jac.sub.",SUBNAME,".RData")
      load(filename)
      dist.used <- jac.sub.dist
      nmds.used <- jac.sub.nmds
      pcoa.used <- jac.sub.pcoa
      rm(jac.sub.dist, jac.sub.nmds, jac.sub.pcoa)
    }
load("R_objects/Bray.RData")
load(params$input)
METRIC <- "Bray"
# Transform original data
phy.ra <- transform_sample_counts(phy, function(x) x/sum(x))
phy.ra
head(otu_table(phy.ra)[,1:6])
bray.dist <- distance(phy.ra, method = "bray",)
# Create matrix from distances
bray.mtx <- melt(as.matrix(bray.dist))
bray.mtx

# Remove self-comparisons
bray.mtx <- bray.mtx %>%
  filter(as.character(Var1) != as.character(Var2)) %>%
  mutate_if(is.factor,as.character)
bray.mtx

# Get sample data
sd <- data.frame(sample_data(phy))
sd <- sd %>%
select (ID, treatment) %>%
mutate_if(is.factor,as.character)

# Combine distances with sample data
colnames(sd) <- c("Var1","Type1")
bray.sd <- left_join(bray.mtx, sd, by = "Var1")

colnames(sd) <- c("Var2","Type2")
bray.sd <- left_join(bray.sd,sd,by = "Var2")

write.csv(file = "output/bdiv_comparison_boxplot.csv",bray.sd)

# Create boxplot
bdiv.comp <- ggplot(bray.sd, aes(x = Type2, y = value)) +
  theme_bw() +
  geom_point() +
  geom_boxplot(aes(color = ifelse(Type1 == Type2, "red", "black"))) +
  scale_color_identity() +
  facet_wrap(~ Type1, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  ggtitle(paste0("Distance metric = ",METRIC))

suppressMessages(ggsave(filename = paste0("plots/bdiv/",SUBNAME,"/",METRIC,"_boxplot_",SUBNAME,".png"), bdiv.comp, device = "png", units = "mm", width = 350, height = 200, dpi = 300))
}

```

# SETTINGS {.tabset .tabset-fade .tabset-pills}

Overview of the parameters and packages that were used for this analysis

## PARAMETERS

The following paramenters were set in for this analysis:

```{r parameters, eval=TRUE}
params <- readRDS("R_objects/bdiv_params.RDS")

tmp <- unlist(params)
dat <- data.frame(Parameter = names(tmp), Value = unname(tmp))


kbl(dat, row.names = F) %>% kable_classic(lightable_options = "striped")

```

## PACKAGES

The analysis was run in R version `r getRversion()` using the following packages:

```{r packages, eval=TRUE}
pack <- data.frame(Package = (.packages()))

for (i in seq(nrow(pack))) pack$Version[i] <- as.character(packageVersion(pack$Package[i]))

kbl(pack[order(pack$Package),], row.names = F) %>% kable_classic(lightable_options = "striped")   
     
```
